---
title: "Clean Multiple Choice Data"
author: "Team X"
date: "`r Sys.Date()`"
output: html_document
---
## import Libraries
Use these libraries to read and clean the data.
```{r}
library(tidyverse)
```

## Load the Data
The data is loaded from the GitHub shared repository.
```{r}
path = 'https://raw.githubusercontent.com/AlphaCurse/DATA607Project3/main/multipleChoiceResponses.csv'
mcr = read.csv(path, na.strings=c("", "NA"))
head(mcr)
```

## Clean the Data
Filter by necessary columns. 
```{r}
mcr1 = mcr %>%
  select(1, 2, 3, 4, 9, 10)
```

Drop duplicated rows.
```{r}
mcr2 = mcr1[!duplicated(mcr), ]
```

Check for missing data.
```{r}
is.na(mcr2)
```
Create function to determine most frequent(mode) values.
```{r}
cal_mode = function(x){
  dist_val = unique(x)#lists distinct/unique values
  dist_tab = tabulate(match(x,dist_val)) #count occurances
  dist_val[which.max(dist_tab)]
}
```

Replace missing data with the mode of the specified column.
```{r}
mcr3 = mcr2 %>%
  mutate(GenderSelect = if_else(is.na(GenderSelect),
                                cal_mode(GenderSelect),
                                GenderSelect))
```

```{r}
mcr4 = mcr3 %>%
  mutate(TitleFit = if_else(is.na(TitleFit),
                                cal_mode(TitleFit),
                                TitleFit))
```

Replace missing data with the median of the specified column.
```{r}
mcr4$Age[is.na(mcr4$Age)] = median(mcr4$Age, na.rm=TRUE)
```

Replace missing data with specified value of the specified column.
```{r}
mcr4$Country = mcr4$Country %>%
  replace_na('Other')
mcr4$CurrentJobTitleSelect = mcr4$CurrentJobTitleSelect %>%
  replace_na('Other')
```

Verify no missing data.
```{r}
summary(mcr4)
```

View data to ensure changes were made.
```{r}
View(mcr4)
```